---
title: Analysis of replicated *Tetranychus* metapopulation experiment - abundance and
  temporal variability
author: "Stefano Masier, Maxime Dahirel, Frederik Mortier, Dries Bonte (this code by M. Dahirel and S. Masier)"
date:
output: 
  html_document:
    theme: yeti
    toc: TRUE
    toc_float: TRUE
editor_options:
  chunk_output_type: console
bibliography: tetranychus-metapop-2019-refs.bib
csl: journal-of-animal-ecology.csl
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, message = FALSE)
```


# A brief introduction

(see manuscript/preprint text for more context)

This script analyses data resulting from the study of experimental metapopulations of spider mites *Tetranychus urticae*. Each replicate metapopulation was composed of 9 patches connected by bridges; the bridge length varied between replicates. Abundances of adult females were tracked weekly in each patch of each metapopulation. We are interested in the effect of bridge length and randomization treatment (see text) on mean patch-level population size, mean metapopulation size and various variability metrics. 

This script relies heavily on de Villemereuil *et al.* [-@villemereuilGeneralMethodsEvolutionary2016] (for converting latent means and variance-covariance matrices to data/observed scale) and Wang and Loreau [@wangEcosystemStabilitySpace2014](for the principle and formula of $\alpha$, $\beta$, $\gamma$ variability in metapopulations). Everything below assumes the reader has read these two papers.

# Part 1 : Preparation

## 1A - packages

First, we’re going to load all the packages we need. This includes the `tidyverse` packages [@wickhamWelcomeTidyverse2019], for data wrangling at the beginning and summarising/plotting at the end, as well as the packages needed to do Bayesian inference and post-processing of Bayesian model results.
This script uses the Stan language [@carpenterStanProbabilisticProgramming2017; @standevelopmentteamRStanInterfaceStan2018] for model fitting (we can use the `rstan` or the `cmdstanr` implementation; see https://mc-stan.org/ for how to install them). The `brms` package [@burknerBrmsPackageBayesian2017] allows you to write a wide range of Stan models using R syntax; it then does the “translation” before fitting. The `QGglmm` package by de Villemereuil *et al.* [-@villemereuilGeneralMethodsEvolutionary2016] is useful to convert latent scale variance-covariance matrices, like a (G)LMM output, to observed scale matrices, needed for our analysis.

```{r packages-loading}
#library(rstan) ## Stan backend
library(cmdstanr) ## Stan backend (another)
library(brms) ## the interface we are using

library(tidyverse)
library(bayesplot)
library(tidybayes)
library(matrixStats)

library(QGglmm) ## this package is needed to help convert variance-covariance matrices from latent to data scale

library(patchwork) #plotting

library(here)

## some useful default settings
# rstan_options(auto_write = TRUE) #for rstan
options(mc.cores = 2) ## reduce/increase depending on cores available
N_chains <- 4
N_warmup <- 400   ## used for publication: 10000 ## 200 is enough for tests
N_iter <- N_warmup + 200  ## recommended for publication-level quality: 2000 iterations post warmup when all chains combined 
## Nwarmup + 200 is probably good enough for tests

```

**Note:** The finished model object will likely be large (at least several 100s Mb) and take some time to fit (a few hours to half a day on the few laptops it has been tested on). If this causes problems, we invite you to try a model with a much smaller `N_warmup` and number of iterations post-warmup (see comments in code chunk above). It should be close enough to convergence, despite warnings, to allow you to get a general idea of the results.


## 1B – data loading and wrangling

```{r data-load}
raw_data <- read_csv(here("data","tetranychus-metapop-2019-dataset.csv"))
```

The variables in `raw_data` are as follow:

- `METAPOP_ID`: unique ID for each replicate metapopulation
- `LENGTH`: length of bridges between patches in the metapopulation (4, 8, or 16 cm)
- `SHUFFLE`: within-metapopulation reshuffling treatment. NO: control, no reshuffling, R: reshuffling
- `REPLICATE`: replicate metapopulation number within a treatment set. The `METAPOP_ID` string corresponds to "`LENGTH` _ `SHUFFLE` _ `REPLICATE`"
- `PATCH`: Patch location in the 3 by 3 metapopulation (written as "row.column")
- `WEEK`: weeks of data collection (first week coded as 1)
- `AFEMA`: number of adult females counted on patch, our measure of abundance

In the raw dataset, data are stored so that there is one column for all population sizes, and a column saying which patch it belongs to. This is a tidy way to store info, but not adequate for what we want to do. To fit our multivariate model, we need one column per patch position instead:

```{r data-reshape}

data <- raw_data %>% 
  mutate(PATCH2= paste("P", PATCH, sep = "")) %>% 
  mutate(PATCH2= str_remove(PATCH2, "[.]")) %>% 
### the two lines above make patches names (a) easier to use as column names (letter as 1st character)
### and consistent with brms standards on response variable names (best to avoid dots and underscore,
### as some functions will remove them for output names and then matching input and output becomes slightly harder)
  select(-c(PATCH)) %>%
  pivot_wider(names_from = PATCH2, values_from = AFEMA) %>%
  drop_na() #3 out of ~550 rows contain NAs in at least one patch, discard


Npatches <- length(unique(raw_data$PATCH)) ### as a check, counting how many patches there are, should be 9
### stored as may be useful later

```

The way the dataset is now, one row = one snapshot of one of the metapopulations (i.e. its state at one time point). We finish by adding a metapopulation total column, for the metapopulation-level model:

```{r metapop-sum}
data <- data %>% 
  mutate(METAPOPSUM = select(.,P11:P33) %>% rowSums())
```


## 1-C Functions for $\alpha$, $\beta$, $\gamma$ variability

One of the aims of this script is to estimate the $\alpha$, $\beta$ and $\gamma$ variability metrics proposed by Wang and Loreau [-@wangEcosystemStabilitySpace2014], while accounting for among metapopulation-variability. The short functions below calculate these metrics, given as input a temporal variance-covariance matrix, and a vector of mean population sizes (both must be numeric and contain the same number of patches, of course).

```{r wang-loreau}

alpha_wang_loreau <- function(varcorr, means) {
  varcorr= as.matrix(varcorr) #ensure there is no problems with diag(varcorr) if varcorr is a 1*1 matrix
  #Basic error checks
    if(!is.vector(means)) stop("Error(not a vector): means should be a vector of numeric values (abundances)")
    if(!is.numeric(means)) stop("Error(not numeric): means should be a vector of numeric values (abundances)")
    if(length(dim(varcorr))>2) stop("Error: varcorr matrix should be a 2-dimensional matrix")
    if(dim(varcorr)[1] != dim(varcorr)[2]) stop("Error: varcorr matrix should be a square matrix")
    if(dim(varcorr)[1] != length(means)) stop("Error: varcorr matrix and means should have same dimensions")
  
  CV_L = sum(sqrt(diag(varcorr))) / sum(means)
  return(CV_L^2)
}

gamma_wang_loreau <- function(varcorr, means) {
  varcorr= as.matrix(varcorr) #ensure there is no problems with diag(varcorr) if varcorr is a 1*1 matrix
  #Basic error checks
    if(!is.vector(means)) stop("Error(not a vector): means should be a vector of numeric values (abundances)")
    if(!is.numeric(means)) stop("Error(not numeric): means should be a vector of numeric values (abundances)")
    if(length(dim(varcorr))>2) stop("Error: varcorr matrix should be a 2-dimensional matrix")
    if(dim(varcorr)[1] != dim(varcorr)[2]) stop("Error: varcorr matrix should be a square matrix")
    if(dim(varcorr)[1] != length(means)) stop("Error: varcorr matrix and means should have same dimensions")
  
  CV_M = sqrt(sum(varcorr)) / sum(means)
  return(CV_M^2)
}

# no function for beta 1, it's simply 1/phi or alpha/gamma
# the beta 2 and the phi functions are here for completeness and checks, we don't need them
# because both betas can be defined through alpha and gamma
# beta 1: see above, beta2 = alpha - gamma

phi_wang_loreau <- function(varcorr) {
  varcorr= as.matrix(varcorr) #ensure there is no problems with diag(varcorr) if varcorr is a 1*1 matrix
  #Basic error checks
  if(length(dim(varcorr))>2) stop("Error: varcorr matrix should be a 2-dimensional matrix")
  if(dim(varcorr)[1] != dim(varcorr)[2]) stop("Error: varcorr matrix should be a square matrix")
    
  return(sum(varcorr) / (sum(sqrt(diag(varcorr))))^2)
}

beta2_wang_loreau <- function(varcorr, means) {
  varcorr= as.matrix(varcorr) #ensure there is no problems with diag(varcorr) if varcorr is a 1*1 matrix
  #Basic error checks
    if(!is.vector(means)) stop("Error(not a vector): means should be a vector of numeric values (abundances)")
    if(!is.numeric(means)) stop("Error(not numeric): means should be a vector of numeric values (abundances)")
    if(length(dim(varcorr))>2) stop("Error: varcorr matrix should be a 2-dimensional matrix")
    if(dim(varcorr)[1] != dim(varcorr)[2]) stop("Error: varcorr matrix should be a square matrix")
    if(dim(varcorr)[1] != length(means)) stop("Error: varcorr matrix and means should have same dimensions")
  
  sqrt_w_bar <- sum(sqrt(diag(varcorr))) / length(means)
  return((sqrt_w_bar^2 - (sum(varcorr)) / (length(means)^2)) / mean(means)^2)
}
```

# Part 2 - Fitting the model

## 2A - Model description

We fit a multivariate generalized linear mixed/multilevel model to the abundance data, with one submodel for each of the 9 patches. This Poisson model (because count data) includes "random" effects of metapopulation ID (to account for the fact some replicates may have higher/lower average population sizes). It also includes a time random effect, to account for temporal patch variance-covariance.

The formula for the model for the number of adult females $N_{i,x,y,t}$ in metapopulation $i$, in the patch of coordinates $x,y$ at time $t$ is

$$\begin{equation*}
N_{[i,x,y,t]} \sim  {\textrm{Poisson}}(\lambda_{[i,x,y,t]}) \\
\log(\lambda_{[i,1,1,t]}) = \alpha_{[1,1]\textrm{TREATMENT}[i]} + \beta_{[i,1,1]} + \gamma_{[i,1,1,t]} \\
...\\
\log(\lambda_{[i,3,3,t]}) = \alpha_{[3,3]\textrm{TREATMENT}[i]} + \beta_{[i,3,3]} + \gamma_{[i,3,3,t]} \\

\mathbf{}\left[\begin{matrix}
\beta_{[i,1,1]} \\ ... \\ \beta_{[i,3,3]} 
\end{matrix}\right] \sim {\textrm{MVNormal}}\left(

\mathbf{}
\left[\begin{matrix}
0 \\ ... \\ 0 
\end{matrix}\right] ,
\boldsymbol{\Omega}_{\textrm{METAPOP}}\right)\\

\left[\begin{matrix}
\gamma_{[i,1,1,t]} \\ ... \\ \gamma_{[i,3,3,t]} 
\end{matrix}\right] \sim {\textrm{MVNormal}}\left(

\left[\begin{matrix}
0 \\ ... \\ 0 
\end{matrix}\right] ,
\boldsymbol{\Omega}_{\textrm{TIME}[i]}\right)\\

\boldsymbol{\Omega}_{\textrm{METAPOP}} = 
\left(\begin{matrix}
\sigma_{M[1,1]}^2 & 0 & ... \\ 0 & \ddots  \\ \vdots & & \sigma_{M[3,3]}^2
\end{matrix}\right)    \\

\boldsymbol{\Omega}_{\textrm{TIME}[i]} = 
\left(\begin{matrix}
\sigma_{T[i,1,1]} & 0 & ... \\ 0 & \ddots  \\ \vdots & & \sigma_{T[i,3,3]} 
\end{matrix}\right)  
\boldsymbol{R}_{\textrm{TIME}[i]}
\left(\begin{matrix}
\sigma_{T[i,1,1]} & 0 & ... \\ 0 & \ddots  \\ \vdots & & \sigma_{T[i,3,3]} 
\end{matrix}\right)    \\

\end{equation*}$$

where $\alpha$ are treatment-specific latent patch intercepts, $\beta$ metapopulation-specific deviations from these intercepts, $\gamma$ temporal abundance fluctuations (not to be confused with $\alpha$, $\beta$, $\gamma$ variabilities), $\boldsymbol{\Omega}$ the relevant covariance matrices and $\boldsymbol{R}_{\textrm{TIME}[i]}$ the temporal correlation matrix for the replicate $i$.

You can see we are estimating a separate temporal variance-covariance matrix for each replicate $i$. We will average everything downstream as needed, but given the non-linearities everywhere in a GLMM, and the fact patch names are arbitrary (we could have rotated metapopulations without changing anything meaningful), it is probably best for the variabilities to calculate everything replicate by replicate first and only average later (we'll come back to these non-linearities later). It is important to note that it means this is a model with observation-level random effects [@harrisonUsingObservationlevelRandom2014], so no further overdispersion to worry about.

Similarly, we can write a (simpler) model for the total metapopulation size $M$ (the `METAPOPSUM` column):

$$
\begin{equation*}
M_{[i,t]} \sim  {\textrm{Poisson}}(\lambda_{[i,t]}) \\
\log(\lambda_{[i,t]}) = \alpha_{\textrm{TREATMENT}[i]} + \beta_{[i]} + \gamma_{[i,t]} \\

\beta_{[i]} \sim \mathrm{Normal}(0, \sigma_{\beta}) \\
\gamma_{[i,t]} \sim \mathrm{Normal}(0, \sigma_{\gamma[i]})

\end{equation*}
$$
(the temporal variance is again replicate-specific and serves again as an observation-level random effect)

## 2B - Implementation - formulas

This is implemented in R as follow. First, for simplicity we create a `TREATMENT` variable combining `LENGTH` and `SHUFFLE`. The `WEEKrpl` variable is useful to fit separate random effects per replicate (this is because in `brms`, random effect levels for each metapop need to have different names if you fit separate random effect matrices per metapopulation). We're also going to create some duplicate columns, just so that the random effects from the patch-by-patch model and from the "all-metapop" model don't end up in the same slots of the fitted model object (it complicates postprocessing otherwise):

```{r treatment-create}
data <- data %>% 
  mutate(TREATMENT = interaction(LENGTH, SHUFFLE)) %>% 
  mutate(WEEKrpl = paste(WEEK, METAPOP_ID)) %>% 
  mutate(WEEKrpl2 = WEEKrpl) %>% 
  mutate(METAPOP_ID2 = METAPOP_ID)
```

Then we build the formula for both the patch by patch multivariate model and the metapopulation total univariate model:

```{r formula}
## see vignette("brms_multivariate") for help on the syntax, in particular the (1|P|random effect) syntax
## see ?gr() for help too if needed
metapopformula <- mvbf(
      bf(P11 ~ 0 + TREATMENT + (1 | METAPOP_ID) + (1 | P | gr(WEEKrpl, by = METAPOP_ID)), family = poisson),
      bf(P12 ~ 0 + TREATMENT + (1 | METAPOP_ID) + (1 | P | gr(WEEKrpl, by = METAPOP_ID)), family = poisson),
      bf(P13 ~ 0 + TREATMENT + (1 | METAPOP_ID) + (1 | P | gr(WEEKrpl, by = METAPOP_ID)), family = poisson),
      bf(P21 ~ 0 + TREATMENT + (1 | METAPOP_ID) + (1 | P | gr(WEEKrpl, by = METAPOP_ID)), family = poisson),
      bf(P22 ~ 0 + TREATMENT + (1 | METAPOP_ID) + (1 | P | gr(WEEKrpl, by = METAPOP_ID)), family = poisson),
      bf(P23 ~ 0 + TREATMENT + (1 | METAPOP_ID) + (1 | P | gr(WEEKrpl, by = METAPOP_ID)), family = poisson),
      bf(P31 ~ 0 + TREATMENT + (1 | METAPOP_ID) + (1 | P | gr(WEEKrpl, by = METAPOP_ID)), family = poisson),
      bf(P32 ~ 0 + TREATMENT + (1 | METAPOP_ID) + (1 | P | gr(WEEKrpl, by = METAPOP_ID)), family = poisson),
      bf(P33 ~ 0 + TREATMENT + (1 | METAPOP_ID) + (1 | P | gr(WEEKrpl, by = METAPOP_ID)), family = poisson),
    
      bf(METAPOPSUM ~ 0 + TREATMENT + (1|METAPOP_ID2) + (1|gr(WEEKrpl2, by = METAPOP_ID2)), family = poisson)
    )
```

note the `~0 + TREATMENT`, rather than `~TREATMENT` syntax in the fixed effects. It’s just a writing trick so we don't need to deal with contrasts and to reduce the quantity of post-processing calculations needed, and also to be sure all prior for latent means have the same precision. See Schielzeth [-@schielzethSimpleMeansImprove2010] and McElreath [-@mcelreathStatisticalRethinkingBayesian2020].  

See also the fact we "combined" the multivariate patch-level model and the univariate metapopulation-level model in one object (they aren't connected though, see the random effects). It's just to fit them in the same `brm` call to make it easier to compare the posteriors later if we want.

## 2C - Implementation - choosing priors

We then select our priors. 

For the fixed effects coefficients (corresponding to latent, i.e. on log scale, grand means of single patch population size), I based myself roughly on the entire distribution of non-zero population densities in De Roissart *et al.* [-@deroissartDataSpatialSpatiotemporal2016; -@deroissartSpatialSpatiotemporalVariation2015] (multiplied to account for our patch size: 5 by 5 cm). Note that the prior we use is slightly wider than what could be done based on density distribution in De Roissart *et al.* (which would be something like $\textrm{Normal}(1.8,0.9)$ for one patch), mostly to have round numbers, but also to allow for a slightly wider prior.  We used a $\textrm{Normal}(4.2,1)$ for the corresponding parameter for the metapopulation-level model; this correspond roughly to a multiplication by 9 of the prior mean compared to the prior chosen for the patch-level approach, and there are 9 patches in each metapopulation.

For the random effects, we started with weakly informative half-normal priors [@mcelreathStatisticalRethinkingBayesian2020]. We placed slightly stronger priors on the temporal variances (observation-level random effects). These are individual patch-specific, so the idea is that if we don't do that, one rare outlier in a patch time series can have an undue influence and drive the variance too high. Given the temporal variances are used to predict the observed-scale means, that would lead to artificially inflated predicted means downstream.

```{r prior-setting}
  prior <- c(
    ### PRIORS FOR THE PATCH-LEVEL MODEL
    set_prior("normal(2,1)", class = "b", 
              resp = c("P11", "P12", "P13", "P21", "P22", "P23", "P31", "P32", "P33")),
    set_prior("normal(0,1)", class = "sd", group = "METAPOP_ID", 
              resp = c("P11", "P12", "P13", "P21", "P22", "P23", "P31", "P32", "P33")),
    set_prior("normal(0,0.5)", class = "sd", group="WEEKrpl", 
              resp = c("P11", "P12", "P13", "P21", "P22", "P23", "P31", "P32", "P33")),
    set_prior("lkj(2)", class = "cor"), # mildly skeptical of very high correlations (positive or negative)
    ### PRIORS FOR THE METAPOP LEVEL MODEL
    set_prior("normal(4.2,1)", class = "b", resp = "METAPOPSUM"),
    set_prior("normal(0,1)", class = "sd", group = "METAPOP_ID2", resp = "METAPOPSUM"),
    set_prior("normal(0,0.5)", class = "sd", group = "WEEKrpl2", resp = "METAPOPSUM")
  )
```

## 2D - Implementation - fitting

We can now fit the model based on the choices we made earlier

```{r model_fitting}
if(file.exists(here("R_output","model.Rdata"))){
  # this if-else statement is avoid re-fitting a model if there is already one existing in R_output
  # to override, re-run the model and re-save manually by selecting only the relevant code lines (or delete the Rdata object before relaunching this code chunk)
  load(here("R_output","model.Rdata"))
  }else
    {

mod <- brm(metapopformula, data = data, 
           iter = N_iter, warmup = N_warmup, chains = N_chains, 
           prior = prior, 
           seed = 42,
           control = list(adapt_delta = 0.9), #default is 0.8; adapt_delta and other controls can be adjusted if needed
           backend= "cmdstanr")

## with rstan as backend, the model *may* throw some some rhats=NA warnings on exit
## they can (for once) be safely ignored: related issue: https://github.com/paul-buerkner/brms/issues/865
## these warnings don't pop up when cmdstanr is used as backend 
## (other "normal" warnings will appear if needed (e.g during a test run with low iters) with both backends)
## or when doing summary(mod)
## you can check that only parameters that should be constant give NA warnings through rhat(mod)
## and quickly look at the rhats:
## hist(rhat(mod));abline(v=1.01,col="red")
save(list="mod", file=here("R_output","model.Rdata"))
}
```

Note that it can take some time to finish (around 2000-4000 iterations/h/core on my laptop).

```{r model_summary}
summary(mod)  #look at Rhat and ESS both tail and bulk
```

You also need to check that the model can reproduce the data well, through posterior predictive checks [@gabryVisualizationBayesianWorkflow2019]
(here, given we have observation-level random effects, it should be OK)

```{r model_ppchecks}

pp_check(mod, resp = "P11", nsamples = 20)
## can the model generate the observed data? (there are many ppcheck possibles, see the help of the bayesplot package)
## in a multivariate model, you need to specify the response (so here which patch)
```

# Part 3 - Obtaining key metrics from model output

## Extraction

Once we have fitted and chosen a model, it's time to extract what we need from it. 

There's going to be a lot of variables, so we're gonna try to follow this naming convention to keep things clear:

- columns or objects prefixed by `P_` refer to the patch by patch model
- ...prefixed by `M_` to the metapop-level model
- names containing `rpl` are for things related to the among-replicates level of variation
- names with `time` for the within-replicate (temporal) level of variation

With that out of the way, let's start by extracting all the fixed effects:

```{r P_M_extract-fixef}
P_M_fixef <- fixef(mod, summary = FALSE) %>% ### all model fixed effects, with one column = one "variable" * treatment combination
  as_tibble() %>%
  mutate(.iteration = 1:dim(.)[1]) %>% ### we put the implicit iteration # into an explicit column
  pivot_longer(-.iteration) %>%        ### we pivot so we can split the "variable*treatment" into patch ID and treatment variables
  mutate(variable=str_extract(name,"P11|P12|P13|P21|P22|P23|P31|P32|P33|METAPOPSUM"),
         LENGTH=as.numeric(str_extract(name,"4|8|16")),
         SHUFFLE=str_extract(name,"NO$|R$")) %>%  # $ : end of string
  select(-name) %>%
  pivot_wider(values_from=value,names_from=variable) %>%  ###we repivot back so that 1 column per patch
  group_by(.iteration, LENGTH, SHUFFLE) %>%  ### we group and nest by iteration (with treatment variables along for the ride)
  nest(P_fixef = c(P11:P33)) %>% 
  mutate(P_fixef = map(.x=P_fixef, .f = ~.x %>% unlist())) %>%   #convert list column from tibbles to vectors for faster calculations
  rename(M_fixef = METAPOPSUM)
```

We end up with a tibble where one row = one posterior iteration of one replicate metapopulation. Where there are more than one value per row per "thing" of interest (like with patch-level fixed effects: 9 patches), we use list-columns to store them.

Let's continue with the among-metapopulation random effects (not the variances):

```{r P_extract-ranef-rpl}
### we then extract the among-metapop deviations for the patch by patch model (needed for estimating alpha beta gamma):
P_ranef_rpl<- ranef(mod, summary = FALSE)$METAPOP_ID %>% 
  array_tree(margin=c(1,2)) %>% ## convert array to list, enlisting by metapop nested in iteration
  tibble() %>% ## we "flatten" the list, iteration become rows of new tibble
  rename(., P_ranef_rpl=".") %>% 
  mutate(.iteration = 1:dim(.)[1]) %>% ### we put the implicit iteration # into an explicit column
  unnest_longer(P_ranef_rpl, indices_to = "METAPOP_ID") %>% 
  mutate(LENGTH=as.numeric(str_extract(METAPOP_ID,"^4|^8|^16")),  ## ^: start of string
         SHUFFLE=str_extract(METAPOP_ID,"NO|R"))
```

and the actual among-metapopulation variances (converted to VCV matrix format):

```{r P_extract-vcv-rpl}
P_vcv_rpl <- VarCorr(mod, summary = FALSE)$METAPOP_ID$sd %>% 
  array_tree(margin=1) %>% 
  tibble() %>% 
  rename(., P_vcv_rpl_latent=".") %>% 
  mutate(.iteration = 1:dim(.)[1]) %>% 
  mutate(P_vcv_rpl_latent=map(.x=P_vcv_rpl_latent,
                       .f=function(.x){diag(.x^2)}))
```


Last thing for the patch level, we need to extract the temporal variance-covariance matrices. We need to do this in two steps. First, we extract the content of the VCV slot:

```{r P_extract-vcv-time1}
memory.limit(40000) #### if needed, a little memory boost to handle the full vcv matrix in final model
P_vcv_time_global <- VarCorr(mod, summary = FALSE)$WEEKrpl$cov
```

Because of the way the model is structured, **all** the variances-covariances that involve `WEEKrpl` are stored in one big matrix, which include across replicate covariances (like patch 1.1 of replicate 1 to patch 1.2 of replicate , for instance _ these are of course 0 by design). We need to split it into its replicate components:

```{r P_extract-vcv-time2}
P_vcv_time_latent<-P_ranef_rpl %>% 
  select(.iteration,METAPOP_ID) %>% 
  mutate(P_vcv_time_latent=map2(.x=.iteration,.y=METAPOP_ID,
                                .f=function(iter=.x,metapop=.y,source=P_vcv_time_global){
                                  include <- colnames(source)
                                  include <- include[str_detect(include, pattern = metapop)]
                                  return(source[iter,include,include])
                                }))

rm(P_vcv_time_global);gc() ##we remove the big global VCV from memory since we don't need it anymore

```


Now we can move to extracting things from the metapopulation size model, where things are a bit simpler.

First the among-replicate random effects:
```{r M_extract-ranef-rpl}
### we do the same for the metapop sum model:
M_ranef_rpl<- ranef(mod, summary = FALSE)$METAPOP_ID2 %>%
  array_tree(margin=c(1)) %>% ## convert array to list, enlisting by metapop nested in iteration
  tibble() %>% ## we "flatten" the list, iteration become rows of new tibble
  rename(., M_ranef_rpl=".") %>% 
  mutate(M_ranef_rpl = map(.x = M_ranef_rpl,
                                     .f = ~ t(.x) %>% as.data.frame() %>% c())) %>% 
  mutate(.iteration = 1:dim(.)[1]) %>% ### we put the implicit iteration # into an explicit column
  unnest_longer(M_ranef_rpl, indices_to = "METAPOP_ID") %>% 
  mutate(LENGTH=as.numeric(str_extract(METAPOP_ID,"^4|^8|^16")),  ## ^: start of string
         SHUFFLE=str_extract(METAPOP_ID,"NO|R"))
```

Then the among-replicate variance:
```{r M_extract-var-rpl}
M_var_rpl <- VarCorr(mod,summary=FALSE)$METAPOP_ID2$sd %>% 
  as.data.frame() %>% 
  mutate(.iteration = 1:dim(.)[1]) %>% 
  mutate(M_var_rpl_latent = METAPOPSUM_Intercept^2) %>% 
  select(.iteration,M_var_rpl_latent)
```

And the replicate by replicate temporal variances:

```{r M_extract-var-time}
M_var_time <- VarCorr(mod,summary=FALSE)$WEEKrpl2$sd %>%
  as.data.frame() %>% 
  mutate(.iteration = 1:dim(.)[1]) %>% 
  pivot_longer(-.iteration) %>% 
  mutate(M_var_time_latent = value^2) %>% 
  mutate(METAPOP_ID = str_remove(name,"METAPOPSUM_Intercept:METAPOP_ID2")) %>% 
  select(METAPOP_ID, M_var_time_latent, .iteration) %>% 
  mutate(LENGTH=as.numeric(str_extract(METAPOP_ID,"^4|^8|^16")),  ## ^: start of string
         SHUFFLE=str_extract(METAPOP_ID,"NO|R"))
```


Now we can merge all of this in one table (still with "one row = one iteration for one replicate metapop"):

```{r merge-coefficients-tables}
tab <- P_ranef_rpl %>% 
  left_join(P_vcv_rpl) %>% 
  left_join(P_vcv_time_latent) %>% 
  left_join(M_ranef_rpl) %>% 
  left_join(M_var_rpl) %>% 
  left_join(M_var_time) %>%  
  left_join(P_M_fixef)
```

## Estimation of key quantities

We first use information on patch-level fixed effects and patch-level temporal variances to estimate data-scale patch-level means. Because we allow within-patch variance to vary among replicates (and so among treatments), we can't infer anything about *mean* differences among treatments from fixed effects only, as the data-scale mean depends on both latent mean and latent within-patch variance [see @villemereuilGeneralMethodsEvolutionary2016]. We need to first estimate patch means from latent means and variances, and then average across patches within replicates:

```{r P_means}
tab <- tab  %>% 
  mutate(P_latent_intercept = map2(
    .x = P_fixef, .y = P_ranef_rpl,
    .f= function(.x,.y){.x + .y}
  )) %>% 
  mutate(P_pred = map2( ## patch by patch average
    .x = P_latent_intercept, .y = P_vcv_time_latent,
    .f = ~ exp(.x + diag(.y)/2)  ## analytic form for the Poisson model, see annex villemereuil
  )) %>% 
  ### we then average within metapops:
  mutate(P_mean_all = map(.x = P_pred, .f = ~mean(.x))) %>% 
  mutate(P_mean_corners = map(.x = P_pred, .f = ~mean(.x[c(1,3,7,9)]))) %>% 
  mutate(P_mean_center = map(.x = P_pred, .f = ~mean(.x[c(5)]))) %>% 
  mutate(P_mean_sides = map(.x = P_pred, .f = ~mean(.x[c(2,4,6,8)]))) %>% 
  unnest(c(P_mean_all, P_mean_corners,P_mean_center,P_mean_sides)) 
```

Now, to estimate $\alpha$, $\beta$ and $\gamma$ variabilities, we need to get the temporal variance-covariance matrices from the latent scale to the data scale [see @villemereuilGeneralMethodsEvolutionary2016]. Ideally we'd use the `QGmv...` functions in the `QGglmm` package, but they become terribly inefficient in memory use once the multivariate model has more than 5 variables (and we have 9). But our 9 variables have the same family, so it's actually very easy to do it manually using the closed form formulas in the source papers!! 

The first step is to estimate a $\psi$ matrix, a diagonal matrix containing for each variable its $\psi$ value. And actually, we've already estimated it, since for a Poisson model, $\psi$ is the predicted mean (i.e. `P_pred` for us). So we just have to diagonalise it. Then we can use it to convert the latent scale temporal VCV to the observed scale, and finally use these VCVs and the predicted means to estimate the $\alpha$, $\beta$ and $\gamma$ variabilities:

```{r P_variabilities}
tab <- tab %>% 
  mutate(P_psi = map(.x = P_pred, .f = function(.x){diag(.x)})) %>% 
  mutate(P_vcv_time_obs = map2(.x = P_psi, .y= P_vcv_time_latent,
                                .f = ~ (.x %*% .y %*% t(.x))
  )) %>% 
  mutate( 
    P_alpha = map2(.x = P_vcv_time_obs, .y = P_pred, .f = ~.x %>% alpha_wang_loreau(varcorr=., means = .y)),
    P_gamma = map2(.x = P_vcv_time_obs, .y = P_pred, .f = ~.x %>% gamma_wang_loreau(varcorr=., means = .y))
  ) %>% 
  unnest(cols=c(P_alpha,P_gamma)) %>% 
  mutate(
    P_beta1 = P_alpha / P_gamma, P_beta2 = P_alpha - P_gamma
  )
```


And we can finish by doing all of the above for the metapopulation-level model (at least that is meaningful to do). Here we can directly use the `QGparams` function, since it uses the closed form formulas we used above:

```{r M_means}
tab<- tab %>% 
  mutate(M_pred = exp(M_fixef + (M_var_rpl_latent + M_var_time_latent)/2)) %>% 
  mutate(M_latent_intercept = M_fixef + M_ranef_rpl) %>% 
  mutate(M_alpha = map2(.x = M_latent_intercept, .y = M_var_time_latent,
                        .f = function(.x,.y){
                          M_obscale = QGparams(mu=.x, var.a = .y, var.p = .y, model="Poisson.log", verbose=FALSE)
                          CV_L = sqrt(M_obscale$var.a.obs)/M_obscale$mean.obs
                          return(CV_L^2)
                          })) %>% 
  unnest(M_alpha)
```


## Averaging across replicates

To conclude, we just need to to get, for each treatment, iteration by iteration, the average of every quantity of interest estimated at the replicate level:

```{r P_summary}
P_summary <- tab %>% 
  group_by(.iteration, SHUFFLE, LENGTH) %>% 
  summarise(P_mean_all = mean(P_mean_all), 
            P_mean_corners = mean(P_mean_corners), 
            P_mean_sides = mean(P_mean_sides), 
            P_mean_center = mean(P_mean_center), 
            P_alpha = mean(P_alpha), 
            P_beta1 = mean(P_beta1), 
            P_beta2 = mean(P_beta2), 
            P_gamma = mean(P_gamma)) %>% 
  ungroup() %>% 
  mutate(SHUFFLE=fct_recode(SHUFFLE, control="NO",randomized="R"))
```


```{r M_summary}
M_summary <- tab %>% 
  group_by(.iteration,SHUFFLE,LENGTH) %>% 
  summarise(M_mean=mean(M_pred),
            M_alpha=mean(M_alpha)) %>% 
  ungroup() %>% 
  mutate(SHUFFLE=fct_recode(SHUFFLE, control="NO",randomized="R"))
```


# Part 4- Figures [BIG WORK IN PROGRESS]

Before doing the figures, we need to create color palettes that'll be used throughout:

```{r figure-palettes}
paletteLENGTH <- c("#D55E00", "#E69F00", "#F0E442") #for figures 2 to 4
paletteLOCAL <-  c("#e66101", "#fdb863", "#5e3c99") #for figure 5
```

And we need to summarise the observed values to display them along th eposterior when relevant:

```{r obs}
obssummary <- data %>% 
  pivot_longer(P11:P33) %>% 
  group_by(SHUFFLE,LENGTH,METAPOP_ID,name) %>%
  summarise(P_mean=mean(value), P_se=plotrix::std.error(value),
            M_mean=mean(METAPOPSUM),M_se=plotrix::std.error(METAPOPSUM)) %>% 
  ungroup() %>% 
  mutate(SHUFFLE=fct_recode(SHUFFLE, control="NO",randomized="R"))

```


## patch level means

```{r P_plot}
P_summary %>% group_by(.iteration,LENGTH, SHUFFLE) %>% summarise(mean = mean(P_mean_all)) %>% 
  ggplot()+
  stat_halfeye(aes(x=mean,y=log(LENGTH,base=4)+0.05, fill=factor(LENGTH)),
               orientation="horizontal",.width=c(0.01,0.95))+
  scale_fill_manual(values=paletteLENGTH)+
  geom_pointinterval(
    data = obssummary, 
    aes(x=P_mean, xmin=P_mean-P_se,xmax=P_mean+P_se,y=log(LENGTH,base=4)-0.05), 
    col="grey40",size = 1.5, alpha=0.5, position=position_jitter(height=0.1,width=NULL))+
  scale_x_continuous("mean patch population size (adult females)")+
  scale_y_continuous("bridge length (cm)", breaks=log(c(4,8,16),base=4), labels=c(4,8,16))+
  facet_wrap(~SHUFFLE)+
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95") +
  theme(legend.position = "none")
```

## metapopulation means

```{r M_plot}
M_summary %>% group_by(.iteration,LENGTH, SHUFFLE) %>% summarise(mean = mean(M_mean)) %>% 
  ggplot()+
  stat_halfeye(aes(x=mean,y=log(LENGTH,base=4)+0.05, fill=factor(LENGTH)),
               orientation="horizontal",.width=c(0.01,0.95))+
  scale_fill_manual(values=paletteLENGTH)+
  geom_pointinterval(
    data = obssummary %>% select(SHUFFLE, LENGTH, M_mean,M_se) %>% distinct(), 
    aes(x=M_mean, xmin=M_mean-M_se,xmax=M_mean+M_se,y=log(LENGTH,base=4)-0.05), 
    col="grey40",size = 1.5, alpha=0.5, position=position_jitter(height=0.1,width=NULL))+
  scale_x_continuous("mean metapopulation size (adult females)")+
  scale_y_continuous("bridge length (cm)", breaks=log(c(4,8,16),base=4), labels=c(4,8,16))+
  facet_wrap(~SHUFFLE)+
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95") +
  theme(legend.position = "none")
```


# References




